Owlifier: Creating OWL-DL Ontologies from Simple Spreadsheet-Based Knowledge Descriptions\tnoteref{t1}

### 2008-12-02 docs/pubs/ISEI6/main.tex

\begin{abstract}
  Discovery and integration of data is important in many ecological
  studies, especially those that concern broad-scale ecological
  questions. Discovery and integration is often a difficult and
  time-consuming task for researchers, in part because informal,
  ambiguous, and sometimes inconsistent terms are used to describe
  data semantics.  Ontologies can help address this problem by
  providing a means to define ecological concepts to more consistently
  annotate, relate, and search for data sets.
  %
  However, unlike in molecular biology or biomedicine, few ontology
  development efforts exist within ecology. Ontology development often
  requires considerable expertise in ontology languages and
  development tools, which is often a barrier for ontology creation in
  ecology. In this paper we address this problem by providing an
  approach for ontology creation that allows ecologists to use common
  spreadsheet tools to describe different aspects of an ontology.  We
  present conventions for creating, relating, and constraining
  concepts through spreadsheets, and provide software tools for
  converting these ontologies into equivalent OWL-DL
  representations. We also consider inverse translations, i.e., to
  convert ontologies represented using OWL-DL into our spreadsheet
  format. Our approach allows large lists of terms to be easily
  related and organized into concept hierarchies, and generally
   provides a more intuitive and natural interface for ontology
  development by ecologists.
\end{abstract}

@@

\myblock{Entity Blocks.} Entity blocks are the primary blocks used to
define ontologies. An entity block introduces new OWL classes and
specifies subclass relationships. Imported classes may also be used
within entity blocks by prefixing class names with namespace labels
(as described above).  Rows containing entity blocks take the form
\begin{itemize}
\item[] 
  \begin{tabular}{|l|l|l|l|l|}\hline
    \textsf{entity} & $c_1$ & $c_2$ & \dots & $c_n$ \\ \hline 
  \end{tabular} \hfill ($n \ge 1$)
\end{itemize}
where each class $c_i$ is asserted in the current ontology to subsume
$c_{i+1}$, for $1 \le i < n$. That is, each $c_i$ in an entity block
induces the description-logic axiom $c_{i+1} \sqsubseteq c_i$.  If
both $c_i$ and $c_{i+1}$ are imported classes, we say that the block
defines an ``articulation'' (i.e., mapping) between the two
classes. The following entity block defines a simple subclass
hierarchy.
\begin{tabbing}
  ~~\textsf{entity} \textsf{PhysicalFeature} 
    \textsf{AquaticPhysicalFeature} \textsf{River}
\end{tabbing}
This block states that Physical Feature, Aquatic Physical Feature, and
River are classes; River is a subclass of Aquatic Physical Feature;
and Aquatic Physical Feature is a subclass of Physical Feature. The
following entity block introduces a new class via an imported class.
\begin{tabbing}
  ~~\textsf{entity} \textsf{sweet:MarineEcosystem} 
    \textsf{IntertidalEcosystem}
\end{tabbing}
This block states that Intertidal Ecosystem is a subclass of the
Marine Ecosystem class imported from the SWEET ontology. Similarly,
assuming ``marine'' denotes an existing ontology of marine ecosystem
concepts, the following entity block defines a simple class
articulation.
\begin{tabbing}
  ~~\textsf{entity} \textsf{sweet:MarineEcosystem} 
    \textsf{marine:DeapSeaEcosystem}
\end{tabbing}
This block states that the Deap Sea Ecosystem class of the marine
ontology is a subclass of the Marine Ecosystem class of the SWEET
ontology.

@@

\myblock{Relationship Blocks.} Relationship blocks define
\emph{required} object properties of classes. An object property
within OWL is a property defined between two class instances. Rows
containing relationship blocks take the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|}\hline \textsf{relationship} & $p$ & $c_1$
    & $c_2$ & \dots & $c_n$ \\ \hline
  \end{tabular} \hfill ($n \ge 2$)
\end{itemize}
where $p$ is an object property and each $c_i$ is a class, for $1 \le
i \le n$.  For every class $c_i$, the relationship block induces the
DL axiom $c_i \sqsubseteq \exists p . c_{i+1}$ stating that each
instance of $c_i$ has, amongst possibly other things, a relationship
through $p$ to some instance of $c_{i+1}$.  For example, the following
block states that the class California Voles live in Grassy Areas.
\begin{tabbing}
  ~~\textsf{relationship} \textsf{livesIn} \textsf{CaliforniaVole} 
    \textsf{GrassyArea}
\end{tabbing}
In some cases, a particular property can apply to a sequence of
classes, and for convenience, each such class can be specified in a
single block. For example, consider the following block.
\begin{tabbing}
  ~~\textsf{relationship} \textsf{directlyBelow} \textsf{Hypolimnion} 
    \textsf{Thermocline} \textsf{Epilimnion} 
\end{tabbing}
This block states that, e.g., within a thermally stratified lake, the
Hypolimnion layer is directly below the Thermocline layer, and the
Thermocline layer is directly below the Epilimnion layer.

\myblock{Transitive Blocks.} Transitive blocks are special cases of
relationship blocks where the object property is asserted to be
transitive. If a property $p$ is declared to be transitive, whenever
$p$ relates an individual $o_1$ to an individual $o_2$, and an
individual $o_2$ to an individual $o_3$, then $p$ is also assumed to
relate $o_1$ to $o_3$. Rows containing transitive blocks take the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|}\hline \textsf{transitive} & $p$ & $c_1$
    & $c_2$ & \dots & $c_n$ \\ \hline
  \end{tabular} \hfill ($n \ge 2$)
\end{itemize}
where $p$ is an object property and each $c_i$ is a class, for $1 \le
i \le n$. The following block is a simple example of a transitive
relationship. 
\begin{tabbing}
  ~~\textsf{transitive} \textsf{hasPart} \textsf{Body}
  \textsf{Head} \textsf{Eye} \textsf{Retina}
\end{tabbing} 
This block states that every Body has a Head as a part, every Head has
an Eye as a part, and every Eye has a Retina as a part. Moreover,
because the part-of property here is defined to be transitive, it is
possible to infer that every Body also has an Eye and a Retina through
the corresponding relationship restrictiosn, i.e., $\textsf{Body}
\sqsubseteq \exists \textsf{hasPart} . \textsf{Head}$, $\textsf{Head}
\sqsubseteq \exists \textsf{hasPart} . \textsf{Eye}$, and
$\textsf{Eye} \sqsubseteq \exists \textsf{hasPart} . \textsf{Retina}$.

\myblock{Inverse Blocks.} Inverse blocks state that two object
properties are inverses of each other. If $p_1$ and $p_2$ are defined
to be inverse properties, whenever $p_1$ relates an individual $o_1$
to an individual $o_2$ then $p_2$ is assumd to relate $o_2$ to $o_1$.
Rows containing inverse blocks take the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|}\hline \textsf{inverse} & $p_1$ & $p_2$
\\ \hline
  \end{tabular}
\end{itemize}
where $p_1$ and $p_2$ are object properties. A common example of
inverse properties are \texttt{partOf} and \texttt{hasPart}.

\myblock{}*** HERE ***

\myblock{Attribute Blocks.} Attribute blocks define required
\emph{datatype} properties of concepts. Unlike object properties, a
datatype property within OWL is a property defined between a class
instance and a literal value (e.g., string, integer, etc.). Rows
containing attribute blocks take the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline \textsf{attribute} & $p$ & $c_1$
    & $c_2$ & \dots & $c_n$ \\ \hline
  \end{tabular} \hfill ($n \ge 1$)
\end{itemize}
where $p$ is a datatype property and $c_i$ is a class, for $1 \le i
\le n$. For each concept $c_i$, the attribute block induces the
description logic axiom $c_i \sqsubseteq \exists p$ stating that each
instance of $c_i$ has, amongst possibly other things, a property $p$
with a literal value. As an example, the following ...

\myblock{Value Blocks.} Value blocks define required datatype property
\emph{constant values} for concepts. A value block has the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline \textsf{value} & $P$ & $V$ & $C_1$
    & $C_2$ & \dots & $C_n$ \\ \hline
  \end{tabular} \hfill ($n \ge 1$)
\end{itemize}
where $P$ is a datatype property, $C_i$ is a concept for $1 \le i \le
n$, and $V$ is a datatype value. For each concept $C_i$, the value
block induces the DL axiom \[C_i \sqsubseteq (V \in P)\] stating that
each instance of $C_i$ has a value $V$ for property $P$.  The value
restrictions stated by value blocks are often used for defining
so-called \emph{value partitions} \cite{co-ode}.

\myblock{Minimum Blocks.} Minimum blocks state the minimum number of
properties $P$ an instance of a concept may have.  Minimum blocks have
the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline \textsf{minimum} & $P$ & $N$ & 
    $C_1$ & $C_2$ & $\dots$ & $C_m$
    \\ \hline
  \end{tabular}
\end{itemize}
where $N$ is the minimum number of properties $P$ that instances of
concept $C_1$ may have to instances of concept $C_2$, $C_2$ to $C_3$,
and so on. A cardinality block induces the DL axiom \[C_i \sqsubseteq
(\le N P.C_{i+1})\] stating that each instance of $C_i$ must be related to
at least $N$ unique instances of $C_{i+1}$ via $P$. For example, the blocks
\begin{quote}
\textsf{minimum} \textsf{hasPart} 1 \textsf{Body} \textsf{Head} \textsf{Nose}\\
\textsf{minimum} \textsf{hasPart} 2 \textsf{Head} \textsf{Eye}
\end{quote}
states that a body has at least one head and at least two eyes.

\myblock{Maximum Blocks.} Maximum blocks state the maximum
number of properties $P$ an instance of a concept may have.
Maximum blocks have the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline \textsf{maximum} & $P$ & $N$ & 
    $C_1$ & $C_2$ & $\dots$ & $C_m$
    \\ \hline
  \end{tabular}
\end{itemize}
where $N$ is the maximum number of properties $P$ that instances of
concept $C_1$ may have to instances of concept $C_2$, $C_2$ to $C_3$,
and so on. A cardinality block induces the DL axiom \[C_i \sqsubseteq
(\ge N P.C_{i+1})\] stating that each instance of $C_i$ may be related to
at most $N$ unique instances of $C_{i+1}$ via $P$. For example, the blocks
\begin{quote}
\textsf{maximum} \textsf{hasPart} 1 \textsf{Body} \textsf{Head} \\
\textsf{maximum} \textsf{hasPart} 2 \textsf{Head} \textsf{Eye}
\end{quote}
states that a body has at least one head and at least two eyes.

\myblock{Sufficient Blocks.} Sufficient blocks state that any instance
having a property $P$ to an instance of a concept $C_2$ is a
sufficient condition for being an instance of a concept $C_1$. A
sufficient block has the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|}\hline \textsf{sufficient} & $C_1$ & $P$ & $C_2$
    \\ \hline
  \end{tabular}
\end{itemize}
where $C_1$ is the target concept (i.e., denoting the concept
definition), $P$ is the sufficient property, and $C_2$ is the
sufficient concept. A sufficient block induces the DL axiom \[C_1
\equiv \exists P.C_2\] Sufficient blocks provide a mechansism to
construct simple class definitions (i.e., classes defined precisely by
other classes), primarily for use with value partitions. [NOTE: these
should be anded together?]

@@

@@

>>>>

\begin{abstract}
  Discovery and integration of data is important in many ecological
  studies, especially those that concern broad-scale ecological
  questions. Discovery and integration is often a difficult and
  time-consuming task for researchers, in part because informal,
  ambiguous, and sometimes inconsistent terms are used to describe
  data semantics.  Ontologies can help address this problem by
  providing a means to define ecological concepts to more consistently
  annotate, relate, and search for data sets.  However, unlike in
  molecular biology or biomedicine, few ontology development efforts
  exist within ecology. Ontology development often requires
  considerable expertise in ontology languages and development tools,
  which is often a barrier for ontology creation in ecology. In this
  paper we address this problem by providing an approach for ontology
  creation that allows ecologists to use common spreadsheet tools to
  describe different aspects of an ontology.  We present conventions
  for creating, relating, and constraining concepts through
  spreadsheets, and provide software tools for converting these
  ontologies into equivalent OWL-DL representations. We also consider
  inverse translations, i.e., to convert ontologies represented using
  OWL-DL into our spreadsheet format. Our approach allows large lists
  of terms to be easily related and organized into concept
  hierarchies, and generally provides a more intuitive and natural
  interface for ontology development by ecologists.
\end{abstract}

@@

\myblock{Entity Blocks.} Entity blocks are the primary blocks used to
define ontologies. An entity block introduces new OWL classes and
specifies subclass relationships. Imported classes may also be used
within entity blocks by prefixing class names with namespace labels
(as described above).  Rows containing entity blocks take the form
\begin{itemize}
\item[] 
  \begin{tabular}{|l|l|l|l|l|}\hline
    \textsf{entity} & $c_1$ & $c_2$ & \dots & $c_n$ \\ \hline 
  \end{tabular} \hfill ($n \ge 1$)
\end{itemize}
where each class $c_i$ is asserted in the current ontology to subsume
$c_{i+1}$, for $1 \le i < n$. That is, each $c_i$ in an entity block
induces the description logic axiom $c_{i+1} \sqsubseteq c_i$.  If
both $c_i$ and $c_{i+1}$ are imported classes, we say that the block
defines an ``articulation'' (i.e., mapping) between the two
classes. The following entity block defines a simple subclass
hierarchy.
\begin{tabbing}
  ~~\textsf{entity} \textsf{PhysicalFeature} 
    \textsf{AquaticPhysicalFeature} \textsf{River}
\end{tabbing}
This block states that Physical Feature, Aquatic Physical Feature, and
River are classes; River is a subclass of Aquatic Physical Feature;
and Aquatic Physical Feature is a subclass of Physical Feature. The
following entity block introduces a new class via an imported class.
\begin{tabbing}
  ~~\textsf{entity} \textsf{sweet:MarineEcosystem} 
    \textsf{IntertidalEcosystem}
\end{tabbing}
This block states that Intertidal Ecosystem is a subclass of the
Marine Ecosystem class imported from the SWEET ontology. Similarly,
assuming ``marine'' denotes an existing ontology of marine ecosystem
concepts, the following entity block defines a simple class
articulation.
\begin{tabbing}
  ~~\textsf{entity} \textsf{sweet:MarineEcosystem} 
    \textsf{marine:DeapSeaEcosystem}
\end{tabbing}
This block states that the Deap Sea Ecosystem class of the marine
ontology is a subclass of the Marine Ecosystem class of the SWEET
ontology.

@@

% \myblock{Property Blocks.} Property blocks define basic object property
% domain and range constraints. , where the domain of an
% object property denotes the set of individuals that can have the
% property, and the range denotes the set of individuals that can be
% related to domain individuals through the property. Rows containing
% relationship blocks take the form 
% \begin{itemize}
% \item[]
%   \begin{tabular}{|l|l|l|l|}\hline \textsf{property} & $p$ & $c_1$
%     & $c_2$ \\ \hline
%   \end{tabular} \hfill ($n \ge 2$)
% \end{itemize}
% where $p$ is the object property, class $c_1$ is the property domain,
% and class $c_2$ is the property range.

\myblock{Relationship Blocks.} Relationship blocks define
\emph{required} object properties of classes.  An object property
within OWL is a property defined between two class instances. Rows
containing relationship blocks take the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|}\hline \textsf{relationship} & $p$ & $c_1$
    & $c_2$ & \dots & $c_n$ \\ \hline
  \end{tabular} \hfill ($n \ge 2$)
\end{itemize}
where $p$ is an object property and each $c$ is a class.  For every
class $c_i$, the relationship block induces the DL axiom $c_i
\sqsubseteq \exists p . c_{i+1}$ stating that each instance of $c_i$
is $p$-related to some instance of $c_{i+1}$, for $1 \le i < n$.  For
example, the following block states that instances of the class
California Voles live in Grassy Areas.
\begin{tabbing}
  ~~\textsf{relationship} \textsf{livesIn} \textsf{CaliforniaVole} 
    \textsf{GrassyArea}
\end{tabbing}
In some cases, a particular property can apply to a sequence of
classes, and for convenience, each such class can be specified in a
single block. For example, consider the following block.
\begin{tabbing}
  ~~\textsf{relationship} \textsf{directlyBelow} \textsf{Hypolimnion} 
    \textsf{Thermocline} \textsf{Epilimnion} 
\end{tabbing}
This block states that, e.g., within a thermally stratified lake, the
Hypolimnion layer is directly below the Thermocline layer, and the
Thermocline layer is directly below the Epilimnion layer.

\myblock{Transitive Blocks.} Transitive blocks are special cases of
relationship blocks where the object property is asserted to be
transitive. If a property $p$ is declared to be transitive, whenever
$p$ relates an individual $o_1$ to an individual $o_2$, and an
individual $o_2$ to an individual $o_3$, then $p$ is also assumed to
relate $o_1$ to $o_3$. Rows containing transitive blocks take the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|}\hline \textsf{transitive} & $p$ & $c_1$
    & $c_2$ & \dots & $c_n$ \\ \hline
  \end{tabular} \hfill ($n \ge 2$)
\end{itemize}
where $p$ is an object property and each $c$ is a class.  The
following block is a simple example of a transitive relationship.
\begin{tabbing}
  ~~\textsf{transitive} \textsf{hasPart} \textsf{Body}
  \textsf{Head} \textsf{Eye} \textsf{Retina}
\end{tabbing} 
This block states that every instance of the class Body has a Head as
a part, every instance of the class Head has an Eye as a part, and
every instance of the class Eye has a Retina as a part. Moreover,
because the \textsf{hasPart} property above is defined to be
transitive, it is possible to infer that every instance of Body also
has an Eye and a Retina as a part through the implied relationship
restrictions $\textsf{Body} \sqsubseteq \exists \textsf{hasPart}
. \textsf{Head}$, $\textsf{Head} \sqsubseteq \exists \textsf{hasPart}
. \textsf{Eye}$, and $\textsf{Eye} \sqsubseteq \exists
\textsf{hasPart} . \textsf{Retina}$.

\myblock{Cardinality Blocks.} Cardinality blocks are also similar to
relationship blocks.  We consider three types of cardinality blocks
for defining minimum, maximum, and exact cardinality restrictions. A
minimum block states the smallest number of properties $p$ to distinct
individuals that an individual of a class may have.  Rows containing
minimum blocks take the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline \textsf{min} & $p$ & $m$ & 
    $c_1$ & $c_2$ & $\dots$ & $c_n$
    \\ \hline 
  \end{tabular} \hfill $(n \ge 2)$
\end{itemize}
where $m$ is the minimum number of properties $p$ that instances of
class $c_i$ must have to instances of concept $c_{i+1}$, for $1 \le i
< n$.  For each class $c_i$, a minimum cardinality block induces the
description logic axiom $c_{i} \sqsubseteq \; (\le m) \; p.c_{i+1}$
stating that each instance of $c_i$ must be $p$-related to at least
$m$ unique instances of $c_{i+1}$. The following two blocks
demonstrate simple minimum cardinality constraints.
\begin{tabbing}
  ~~\textsf{min} \textsf{hasPart} 1 \textsf{Body} \textsf{Head} 
  \textsf{Nose} \\
  ~~\textsf{min} \textsf{hasPart} 2 \textsf{Head} \textsf{Eye}
\end{tabbing}
The first block states that instances of the class Body have at least
one Head as a part, which in turn have at least one Nose as a
part.\footnote{Cardinality restrictions ensuring participation to at
  least one property are typically not given through minimum
  cardinality blocks since they are also implied by relationship
  blocks.} The second block states that instances of the class Head
have at least two Eyes as parts.

A maximum block states the largest number of properties $p$ to
distinct individuals that an individual of a class may have.  Rows
containing maximum blocks take the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline \textsf{max} & $p$ & $m$ & 
    $c_1$ & $c_2$ & $\dots$ & $c_n$
    \\ \hline
  \end{tabular} \hfill $(n \ge 2)$
\end{itemize}
where $m$ is the maximum number of properties $p$ that instances of
concept $c_i$ may have to instances of concept $c_{i+1}$, for $1 \le i
< n$.  For each class $c_i$, a maximum cardinality block induces the
description logic axiom $c_i \sqsubseteq \; (\ge m) \; p.c_{i+1}$
stating that each instance of $c_i$ may be $p$-related to at most $m$
unique instances of $c_{i+1}$. The following two blocks demonstrate
simple maximum cardinality constraints.
\begin{tabbing}
  ~~\textsf{max} \textsf{hasPart} 1 \textsf{Body} \textsf{Head} 
  \textsf{Nose} \\
  ~~\textsf{max} \textsf{hasPart} 2 \textsf{Head} \textsf{Eye}
\end{tabbing}
The first block states that instances of the class Body have at most
one Head as a part, which in turn has at most one Nose as a part. The
second block states that instances of the class Head have at most two
Eyes as parts.

An exact block ensures both a minimum and maximum number $m$ of
properties $p$ to distinct individuals that an individual of a class
must have. Rows containing exact blocks take the form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|l|}\hline \textsf{exact} & $p$ & $m$ & 
    $c_1$ & $c_2$ & $\dots$ & $c_n$
    \\ \hline
  \end{tabular} \hfill $(n \ge 2)$
\end{itemize}
where $m$ is the number of properties $p$ that instances of concept
$c_i$ must have to instances of concept $c_{i+1}$, for $1 \le i < n$.
For each class $c_i$, an exactly block induces the description logic
axiom $c_i \sqsubseteq \; (= m) \; p.c_{i+1}$ stating that each
instance of $c_i$ must be $p$-related to at $m$ unique instances of
$c_{i+1}$.

\myblock{Inverse Blocks.} Inverse blocks state that two object
properties are inverses of each other. If $p_1$ and $p_2$ are defined
to be inverse properties, whenever $p_1$ relates an individual $o_1$
to an individual $o_2$ then $p_2$ (as the inverse of $p_1$) is assumed
to relate $o_2$ to $o_1$.  Rows containing inverse blocks take the
form
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|}\hline \textsf{inverse} & $p_1$ & $p_2$
\\ \hline
  \end{tabular}
\end{itemize}
where $p_1$ and $p_2$ are object properties. A common example of
inverse properties are \texttt{hasPart} and \texttt{partOf}, i.e., if
an individual $o_1$ has an individual $o_2$ as a part, then $o_2$ is
by definition a part of $o_1$.

% \myblock{Attribute Blocks.} Attribute blocks define required
% \emph{datatype} properties of classes. Unlike object properties, a
% datatype property within OWL is a property defined between a class
% instance and a literal value (e.g., a string or integer). Rows
% containing attribute blocks take the form
% \begin{itemize}
% \item[]
%   \begin{tabular}{|l|l|l|l|l|l|l|}\hline \textsf{attribute} & $p$ & $c_1$
%     & $c_2$ & \dots & $c_n$ \\ \hline
%   \end{tabular} \hfill ($n \ge 1$)
% \end{itemize}
% where $p$ is a datatype property and $c_i$ is a class, for $1 \le i
% \le n$. For each concept $c_i$, the attribute block induces the
% description logic axiom $c_i \sqsubseteq \exists p$ stating that
% each instance of $c_i$ has, amongst possibly other things, a
% property $p$ with a literal value. Attributes are often useful for
% As an example, the following ...

% \myblock{Value Blocks.} Value blocks define required datatype property
% \emph{constant values} for concepts. A value block has the form
% \begin{itemize}
% \item[]
%   \begin{tabular}{|l|l|l|l|l|l|l|}\hline \textsf{value} & $P$ & $V$ & $C_1$
%     & $C_2$ & \dots & $C_n$ \\ \hline
%   \end{tabular} \hfill ($n \ge 1$)
% \end{itemize}
% where $P$ is a datatype property, $C_i$ is a concept for $1 \le i \le
% n$, and $V$ is a datatype value. For each concept $C_i$, the value
% block induces the DL axiom \[C_i \sqsubseteq (V \in P)\] stating that
% each instance of $C_i$ has a value $V$ for property $P$.  The value
% restrictions stated by value blocks are often used for defining
% so-called \emph{value partitions} \cite{co-ode}.

\myblock{Sufficient Blocks.} Sufficient blocks are similar to synonym
blocks in that they state equivalences between classes. However,
equivalence relationships are defined in sufficient blocks through
property restrictions (i.e., via anonymous classes). We consider two
main types of sufficient blocks. Rows containing sufficient blocks
take one of the following forms.
\begin{itemize}
\item[]
  \begin{tabular}{|l|l|l|l|l|l|}\hline 
    \textsf{sufficient} & \textsf{min}, \textsf{max}, or 
    \textsf{exact} & $m$ & $c_1$ & $p$ & $c_2$\\ \hline
  \end{tabular}
\item[]
  \begin{tabular}{|l|l|l|l|l|}\hline 
    \textsf{sufficient} & \textsf{only} & $c_1$ & $p$ & \{\textsf{not}\} $c_2$\\ \hline
  \end{tabular}
\end{itemize}
The first form states that individuals that are $p$-related to at
least, at most, or exactly $m$ individuals of class $c_2$ are by
definition individuals of the class $c_1$. Sufficient blocks of this
form induce either the description logic axiom $c_1 \equiv \; (\le m)
\; p . c_2$, $c_1 \equiv \; (\ge m) \; p . c_2$, or $c_1 \equiv \; (=
m) \; p . c_2$ depending on whether the qualifier \textsf{min},
\textsf{max}, or \textsf{exact} is used, respectively. The second form
states that individuals that are $p$-related to only individuals of
class $c_2$ are by definition individuals of the class
$c_1$. Sufficient blocks of this form induce description logic axioms
$c_1 \equiv \forall p . c_2$. The complement of $c_2$ may also be used
within these blocks via the expression \textsf{not $c_2$}. Sufficient
blocks of this form induce description logic axioms $c_1 \equiv
\forall p . \neg c_2$ stating that individuals that are $p$-related to
only individual that are not of class $c_2$ (i.e., the complement of
$c_2$) are by definition individuals of class $c_1$.  Multiple
sufficient blocks for a class $c_1$ are treated as a single
description logic axiom, e.g., if each of the above blocks are defined
for $c_1$ we have $c_1 \equiv \; (\le m \; p .  c_2) \sqcap (\ge m \;
p . c_2) \sqcap (\forall p.c_2)$. The following blocks are simple
examples of sufficient conditions.
\begin{tabbing}
  ~~\textsf{sufficient} \textsf{mimimum} 1 \textsf{MammalProper} 
  \textsf{hasPart} \textsf{MammalProperEar}
  ~~\textsf{sufficient} \textsf{maximum} 2 \textsf{MammalProper} 
  \textsf{hasPart} \textsf{MammalProperEar}
  ~~\textsf{sufficient} \textsf{mimimum} 1 \textsf{MammalProper} 
  \textsf{hasPart} \textsf{Hair}

  ~~\textsf{sufficient} \textsf{mimimum} 3 \textsf{MammalProper} 
  \textsf{hasPart} \textsf{MiddleEarBone}
\end{tabbing}

\begin{tabbing}
  ~~\textsf{sufficient} \textsf{relationship} \textsf{MammalProper} \textsf{hasPart} \textsf{SweatGland} \\
  ~~\textsf{sufficient} \textsf{MammalProper} \textsf{hasPart} \textsf{Hair}
\end{tabbing}
These blocks state that

@@

*** Using owlifier to represent values, value types, and untangling
from owl-pizzas ***

@@

related work: similar to dave's work on ``global taxonomic
constraints'', which assume certain structures of hierarchies, e.g.,
disjoint siblings.
